### 2024-01-25 technical_descriptions.qmd
just a random projection of the training data

---
format:
  pdf:
    keep-tex: true
bibliography: biblio.bib
---

## Example: Random Projections of World Models

- Take the [world_place.csv]( https://github.com/wesg52/world-models/blob/main/data/entity_datasets/world_place.csv) that was used in @gurnee2023language, which maps locations/areas to their latitude and longitude. For each place, it also indicates the corresponding country.
- Take the subset that contains countries that are currently part of the top 10 [FIFA world ranking](https://www.fifa.com/fifa-world-ranking/men?dateId=id14142).
- Assign the current rank to each country, i.e. Argentina gets 1, France gets 2, ...
- Encode all features excluding the FIFA world rank indicator but including `longitude` and `latitude` as continuous variables: $X^{(n \times d)}$ where $n$ is the number of samples and $d$ is the number of resulting features.
- Initialize a small neural network, let's call it a *projector*, that maps from $X$ to a single small hidden layer with sigmoid activation and then from there to a lower-dimensional output space $\mathcal{Z} \subset \mathbb{R}$.
- Without performing any training on the *projector*, simply compute a forward pass of the continuously encoded data $X$ through the *projector* and store the output: $\mathbf{Z}^{(n \times k)}$ where $k$ is the dimension of the latent space of the VAE.
- Next, we perform the linear probe on $\mathbf{Z}$ through Ridge regression: $\mathbf{W} = (\mathbf{Z}'\mathbf{Z} + \lambda \mathbf{I}) (\mathbf{Z}'\mathbf{Y})^{-1}$ where $\mathbf{Y}$ is the $(n \times 2)$ matrix containing the longitude and latitude for each sample. 
- Finally, compute the predicted coordinate for each samples as $\widehat{\mathbf{Y}}=\mathbf{Z}\mathbf{W}$
- Plot the results on a world map and be amazed.

> We can draw a very interesting parallel to @gurnee2023language: we know that geographical coordinates are in the training corpus of Llama-2; that information gets passed through a massive latent space (the model); they run probes on a subsapce of the model (still massive) and manage to linearly separate their variable of interest (coordinates). I don't know how accurate that parallel really is but it just seems striking to me that ultimately what they is regressing a 2-dimensional output variable on this massive feature matrix ("ranging from 4,096 to 8,192 features" for roughly 40,000 total samples, can't figure out how big the train/test sets are, respectively.).

Relatedly, it is worth pointing out that

- a recent paper has shown that the double-descent phenomenon is not exclusive to deep learning but also applies to linear regression with many features [LINK].

## Example: PCA

## Example: Autoencoder

## Example: LLM

## References